---
title: "Multimodal Data Fusion in 3D Printing Quality Prediction"
---
This project presents a novel approach to predict the quality of 3D-printed objects using multiple measurement data sources. Dimension, profile, and surface roughness are represented in image data taken by two sensors of different accuracy and efficiency. Dimension reduction techniques are employed for extracting features from the measurement images. Quality metric is defined fusing profile deviation and surface roughness. The final result is a two-level classification model trained with printing input parameters and measurement data from the two sensors respectively at each stage to predict quality. The model proposes using a preliminary classifier for an initial inspection, followed by a more refined classifier trained by high resolution measurements to further classify the samples predicted as uncertain in the previous round. Such method could guarantee time efficiency while maintaining high accuracy. The result shows feature extraction from high-dimensional image data as a promising technique for efficient and automated quality inspection. 

```{r}
#3D scanner
scanner_folder <- "3D scanner/3d-printing/"
scanner_file_list <- list.files(path = scanner_folder, pattern = "*.csv")
scanner_data_list <- list()
z_list <- list()
for (i in 1:length(scanner_file_list)){
  temp <- assign(scanner_file_list[i], read.csv(paste(scanner_folder, scanner_file_list[i], sep=''), skip = 3, header = FALSE))
  X <- as.numeric(levels(temp$V2))[temp$V2]
  Y <- as.numeric(levels(temp$V3))[temp$V3]
  Z <- as.numeric(levels(temp$V4))[temp$V4]
  z_list[[i]] <- as.numeric(levels(temp[which(X < 35 & Y < 35),]$V4))[temp[which(X < 35 & Y < 35),]$V4]
  z_list[[i]][is.na(z_list[[i]])] <- 0
}
for(i in 1:length(z_list)){
  print(length(z_list[[i]]))
}


z_list[c(4,17,22)] <- NULL

#training and test
training_index <- sample(1:length(z_list), 0.7*length(z_list))
training_set <- data_list[training_index]


scanner_training_set <- z_list[training_index]
scanner_testing_set <- z_list[-training_index]
scanner_matrix <- matrix(c(scanner_training_set[[1]], scanner_training_set[[2]], 
                           scanner_training_set[[3]], scanner_training_set[[4]], 
                           scanner_training_set[[5]], scanner_training_set[[6]], 
                           scanner_training_set[[7]], scanner_training_set[[8]],
                           scanner_training_set[[9]], scanner_training_set[[10]],
                           scanner_training_set[[11]], scanner_training_set[[12]], 
                           scanner_training_set[[13]], scanner_training_set[[14]]),
                           nrow = 14, ncol= 50000)
scanner_pca <- prcomp(scanner_matrix)
summary(scanner_pca)
SPC1 <- scanner_pca$x[,1]
SPC2 <- scanner_pca$x[,2]
SPC3 <- scanner_pca$x[,3]
SPC4 <- scanner_pca$x[,4]

scanner_test_matrix <- matrix(c(scanner_testing_set[[1]], scanner_testing_set[[2]], 
                           scanner_testing_set[[3]], scanner_testing_set[[4]], 
                           scanner_testing_set[[5]], scanner_testing_set[[6]]),
                           nrow = 6, ncol= 50000)
scanner_test_pca <- prcomp(scanner_test_matrix)
summary(scanner_test_pca)
SPC1_test <- scanner_test_pca$x[,1]
SPC2_test <- scanner_test_pca$x[,2]
SPC3_test <- scanner_test_pca$x[,3]
SPC4_test <- scanner_test_pca$x[,4]


```

```{r}
#roughness
roughness_folder <- "roughness3/"
roughness_list <- list.files(path = roughness_folder, pattern = "*.csv")
roughness <- list()

for (i in 1:length(roughness_list)){
  temp_roughness <- data.frame(assign(roughness_list[i], read.csv(paste(roughness_folder, roughness_list[i], sep=''), skip = 7, header = FALSE)))
  area1 <- temp_roughness[1,6]
  area2 <- temp_roughness[2,6]
  roughness[[i]] <- (area1 + area2)/2
}

roughness <- c(roughness[[1]], roughness[[2]],roughness[[3]],
               roughness[[5]], roughness[[6]], roughness[[7]], roughness[[8]], roughness[[9]], 
               roughness[[10]], roughness[[11]], roughness[[12]], roughness[[13]], roughness[[14]],                    roughness[[15]], roughness[[16]], roughness[[17]], roughness[[18]], roughness[[19]],                   roughness[[20]],roughness[[21]])

roughness_training <- roughness[training_index]
roughness_testing <- roughness[-training_index]
```

```{r}
#profile deviation
profile_file <- data.frame(read.csv("profile_quality.csv", header = TRUE))
profile_resid <- profile_file[,2]
profile_training <- profile_resid[training_index]
profile_testing <- profile_resid[-training_index]
```



```{r}
#preliminary regression
model_df <- data.frame(SPC1, SPC2, SPC3, SPC4,
                       c(100,20,40, 40, 40, 60, 60, 60, 
                         100, 100, 20, 20, 40, 60), 
                       c(0.1, 0.1, 0.1, 0.1, 0.1,0.1,0.1, 0.1, 
                         0.2, 0.2, 0.2, 0.2, 0.2, 0.2), 
                       c(20, 40, 20, 40, 60, 20, 40, 60, 40,
                         60, 20, 60, 20, 60))
colnames(model_df) <- c("SPC1", "SPC2", "SPC3", "SPC4","fill", "thickness", "speed")
#model_df$quality <- factor(model_df$quality, levels = sort(unique(model_df$quality)))
attach(model_df)


```

```{r}
#profile
profile_model1 <- lm(profile_training  ~ SPC1 + SPC2+ SPC3 + SPC4 + fill + thickness + speed, data = model_df)
summary(profile_model1)
step_b <- step(profile_model1, direction = "backward")
profile_model1 <- lm(profile_training  ~ SPC1 + SPC2 + SPC4 + fill + thickness, data = model_df)
summary(profile_model1)
#ggpairs(model_df, columns = 1:ncol(model_df), title = "",  axisLabels = "show")

residplot_profile <- resid(profile_model1)
#plot(model_df$SPC4, residplot_profile)
#abline(0,0)

model_df_test_profile <- data.frame(SPC1_test, SPC2_test, SPC4_test,
                            c(100, 100, 20, 100, 40, 40),
                            c(0.1, 0.1, 0.1, 0.2, 0.2, 0.2),
                            c(40, 60, 60, 20, 40, 60))
colnames(model_df_test_profile) <- c("SPC1", "SPC2", "SPC4","fill", "thickness")
MSPE_profile <- mean((profile_testing - predict.lm(profile_model1, model_df_test_profile))^2)
print(MSPE_profile)
predict_profile <- predict.lm(profile_model1, model_df_test_profile)
```

```{r}
#roughness
roughness_model1 <- lm(roughness_training~SPC1 + SPC2 + SPC3 + SPC4 + fill + thickness + speed, data = model_df)
summary(roughness_model1)
#step_b <- step(roughness_model1, direction = "backward")
roughness_model1 <- lm(roughness_training ~ SPC1 + SPC3 + fill + thickness + fill*thickness, data = model_df)
summary(roughness_model1)

model_df_test_roughness <- data.frame(SPC1_test, SPC3_test,
                            c(100, 100, 20, 100, 40, 40),
                            c(0.1, 0.1, 0.1, 0.2, 0.2, 0.2),
                            c(40, 60, 60, 20, 40, 60))
colnames(model_df_test_roughness) <- c("SPC1", "SPC3", "fill", "thickness", "speed")
MSPE_roughness <- mean((roughness_testing - predict.lm(roughness_model1, model_df_test_roughness))^2)
print(MSPE_roughness)
predict_roughness <- predict.lm(roughness_model1, model_df_test_roughness)
```

```{r}
#cluster


name <- c("0.110020", "0.110040", "0.110060", "0.12040", "0.12060", "0.14020", "0.14040", 
          "0.14060", "0.16020", "0.16040", "0.16060", "0.210020", "0.210040", "0.210060",
          "0.22020", "0.22060", "0.24020", "0.24040", "0.24060", "0.26060")

cluster_df <- data.frame(profile_training, roughness_training)
#set.seed(4747)
cluster_kmeans <- kmeans(cluster_df, 3)
plot(cluster_df, col = cluster_kmeans$cluster, pch = 19, main = "Clustering between roughness and profile residual", xlab = "profile residual(um^2)", ylab = "roughness(%)")
points(cluster_kmeans$centers, col = 1:3 , pch = 18, cex = 2)
plot(data.frame(predict_profile, predict_roughness), col = "purple")
legend("topright", legend = c("Good", "uncertain", "bad", "Good mean", "uncertain mean", "bad mean"), col = c("red", "black", "green","red", "black","green"), pch = c(19, 19, 19, 18, 18, 18))
cluster_kmeans$cluster

```



```{r}
#high resolution
folder <- "matrix/"
file_list <- list.files(path = folder, pattern = "*.csv")
data_list <- list()
matrix_list <- list()
for (i in 1:length(file_list)){
  temp <- assign(file_list[i], read.csv(paste(folder, file_list[i], sep=''), skip = 23, header = FALSE))
  temp_matrix <- as.matrix(temp)
  colnames(temp_matrix) <- NULL
  temp_matrix[is.na(temp_matrix)] <- 0
  if(i == 6){
    temp_matrix = t(temp_matrix)
  }
  if(ncol(temp_matrix) == 1634){
    temp_matrix<-temp_matrix[, -c(1629:1634)]
  }
  if(ncol(temp_matrix) == 1648){
    temp_matrix<-temp_matrix[, -c(1629:1648)]
  }
  #print(ncol(temp_matrix))
  matrix_list[[i]] <- temp_matrix
  temp_vector <- as.vector(temp_matrix)
  data_list[[i]] <- temp_vector
}
matrix_list[[4]] <- NULL
```


```{r}
#UMPCA

umpcaTensor2 <-matrix(matrix_list[c(1,4,6,7,8,9,10, 11,13,14, 15,16, 17, 20)])
umpcaMatrix2 <-abind(umpcaTensor2[[1]], umpcaTensor2[[2]], umpcaTensor2[[3]],
                    umpcaTensor2[[4]], umpcaTensor2[[5]], umpcaTensor2[[6]],
                    umpcaTensor2[[7]], umpcaTensor2[[8]], umpcaTensor2[[9]], umpcaTensor2[[10]],
                    umpcaTensor2[[11]], umpcaTensor2[[12]], umpcaTensor2[[13]],
                    umpcaTensor2[[14]],along = 0)
dim(umpcaMatrix2) <- c(1520, 1628, 14)

umpca2 <- UMPCA(umpcaMatrix2, numP = 10)
#umpcaMatrix <- umpcaMatrix - matrix(t(matrix(umpca$TXmean, dim(umpca$TXmean)[1], dim(umpca$TXmean)[2]*numSpl)),dim(umpca$TXmean)[1]*c(1,1),dim(umpca$TXmean)[2]*numSpl, byrow = T)

#centering
tx2 <- umpca2$TXmean
umpcaMatrix2 <- umpcaMatrix2 - abind(tx, tx, tx, tx, tx, tx, tx, tx, tx, tx, tx, tx, tx, tx)
N <- 2
numSpl <- 14
numP <- 10
newfea2 <- matrix(c(0,0,0,0,0,0,0,0,0,0), c(0,0,0,0,0,0,0,0,0,0,0,0,0,0), nrow = 14, ncol = 10)
for (i in 1:numP){
  projFtr2<- MFPCA::ttv(umpcaMatrix2, list(umpca2$Us[[1]][,i],umpca2$Us[[2]][,i]), dim = c(1,2))
  newfea2[, i] <- projFtr2
}
newfea2 <- newfea2[, umpca2$odrIdx]


test_umpcaTensor <- matrix(matrix_list[c(3,5,19)])
test_umpcaMatrix <-abind(test_umpcaTensor[[1]], test_umpcaTensor[[2]], test_umpcaTensor[[3]], along = 0)
dim(test_umpcaMatrix) <- c(1520, 1628, 3)
test_umpca <- UMPCA(test_umpcaMatrix, numP = 3)
#centering
test_tx <- test_umpca$TXmean
test_umpcaMatrix <- test_umpcaMatrix - abind(test_tx, test_tx, test_tx)

test_numP <- 3
test_newfea <- matrix(c(0,0,0), c(0,0,0), nrow = 3, ncol = 3)
for (i in 1:test_numP){
  test_projFtr<- MFPCA::ttv(test_umpcaMatrix, list(test_umpca$Us[[1]][,i],test_umpca$Us[[2]][,i]), dim = c(1,2))
  test_newfea[, i] <- test_projFtr
}
test_newfea <- test_newfea[, test_umpca$odrIdx]

MPC1 <- newfea2[,1]
MPC2 <- newfea2[,2]
MPC3 <- newfea2[,3]
MPC4 <- newfea2[,4]

test_MPC1 <- test_newfea[,1]
test_MPC2 <- test_newfea[,2]
test_MPC3 <- test_newfea[,3]



```


```{r}
#refined regression
profile_highres <- profile_resid[c(1,8,9,14,15,17,20)]
profile_highres2 <- profile_resid[c(1,4,6,7,8,9,10, 11,13,14, 15,16, 17, 20)]
roughness_highres  <- roughness[c(1,8,9,14,15,17,20)]
roughness_highres2  <- roughness[c(1,4,6,7,8,9,10, 11,13,14, 15,16, 17, 20)]
MPCAmodel_df <- data.frame(MPC1, MPC2, MPC3, MPC4,
                       c(100, 40, 60, 100, 20, 40, 60), 
                       c(0.1, 0.1, 0.1, 0.2, 0.2, 0.2, 0.2), 
                       c(20, 60, 20, 60, 20,20, 60))
MPCAmodel_df2 <- data.frame(MPC1, MPC2, MPC3, MPC4,
                            c(100,20,40, 40, 40, 60, 60, 60, 
                         100, 100, 20, 20, 40, 60), 
                       c(0.1, 0.1, 0.1, 0.1, 0.1,0.1,0.1, 0.1, 
                         0.2, 0.2, 0.2, 0.2, 0.2, 0.2), 
                       c(20, 40, 20, 40, 60, 20, 40, 60, 40,
                         60, 20, 60, 20, 60))
colnames(MPCAmodel_df2) <- c("MPC1", "MPC2", "MPC3", "MPC4", "Mfill", "Mthickness", "Mspeed")


```



```{r}
#profile2
profile_model2 <- lm(profile_highres2  ~ MPC1 +Mfill + Mspeed, data = MPCAmodel_df2)
summary(profile_model2)
step_b <- step(profile_model2, direction = "backward")

MPCAmodel_df_test_profile <- data.frame(test_MPC1,
                       c(100, 100, 20, 100, 40, 40),
                       c(40, 60, 60, 20, 40, 60))
colnames(MPCAmodel_df_test_profile) <- c("MPC1", "Mfill", "Mspeed")
profile_testing2 <- profile_resid[c(3,5,19)]
MSPE_profile2 <- mean((profile_testing2 - predict.lm(profile_model2, MPCAmodel_df_test_profile))^2)
print(MSPE_profile2)
```

```{r}
#roughness2
roughness_model2 <- lm(roughness_highres2 ~ MPC3 + Mfill + Mthickness + Mfill*Mthickness, data = MPCAmodel_df2)
summary(roughness_model2)
step_b <- step(roughness_model2, direction = "backward")

MPCAmodel_df_test_roughness <- data.frame(test_MPC3,
                       c(100, 100, 20, 100, 40, 40),
                       c(0.1, 0.1, 0.1, 0.2, 0.2, 0.2),
                       c(40, 60, 60, 20, 40, 60))
colnames(MPCAmodel_df_test_roughness) <- c("MPC3", "Mfill", "Mthickness", "Mspeed")
roughness_testing2 <- roughness[c(3,5,19)]
MSPE_roughness2 <- mean((roughness_testing2 - predict.lm(roughness_model2, MPCAmodel_df_test_roughness))^2)
print(MSPE_roughness2)
```

